Trait coverage for plant data

### 2023-11-01 code/L1/L1_F2_TropicalAndes_frugivore_traits.Rmd

---
title: "Frugivoria frugivore trait subset"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Subsetting Frugivoria to the species and traits of interest"
data input: "TropicalAndes_Frugivoria_frugivore_traits.csv", "TropicalAndes_GBIF_frugivore_occ_cleaned.csv"
data output: ""
date: "2023-08-03"
output: html_document
---

@@

```{r}
source(file.path(data_path_L0, "data_summary.R"))
```

@@

# Write data to csv
```{r}
write.csv(Frugivoria_subset, file.path(output_path,"TropicalAndes_Frugivoria_traits_Forest.csv"))
```

>>>>

---
title: "Frugivoria frugivore trait subset"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Subsetting Frugivoria to the species and traits of interest"
data input: "TropicalAndes_Frugivoria_frugivore_traits.csv", "TropicalAndes_GBIF_frugivore_occ_cleaned.csv"
data output: "TropicalAndes_Frugivoria_traits_subset.csv"
date: "2023-10-31"
output: html_document
---

@@

```{r}
source("~/GitHub/neotropical_plants/code/L0/data_summary.R")
```

@@

# Write data to csv
```{r}
write.csv(Frugivoria_subset, file.path(output_path,"TropicalAndes_Frugivoria_traits_subset.csv"))
```

### 2023-11-01 code/L1/L1_P1_TropicalAndes_plants_GBIF.Rmd

```{r}
table(TropicalAndes_GBIF_plant_occ_cleaned$year)
```

```{r}
# remove records from before second world war
TropicalAndes_GBIF_plant_occ_cleaned <- TropicalAndes_GBIF_plant_occ_cleaned %>%
  filter(year > 1945) 
```

@@

```{r}
print("The number of records is")
nrow(TropicalAndes_GBIF_plant_occ_cleaned)
print("The number of species is")
length(unique(TropicalAndes_GBIF_plant_occ_cleaned$species))
print("The number of families is")
length(unique(TropicalAndes_GBIF_plant_occ_cleaned$family))
```

>>>>

@@

```{r}
print("The number of records is")
nrow(TropicalAndes_GBIF_plant_occ_cleaned)
print("The number of species is")
length(unique(TropicalAndes_GBIF_plant_occ_cleaned$species))
print("The number of genera is" )
length(unique(TropicalAndes_GBIF_plant_occ_cleaned$genus))
print("The number of families is")
length(unique(TropicalAndes_GBIF_plant_occ_cleaned$family))
```

### 2023-11-01 code/L2/L2_TropicalAndes_plant_taxcoverage.Rmd

# Load required packages
```{r}
library(WorldFlora)
library(taxize)
library(BIEN)
library(GIFT)
library(dplyr)
```

# Read in Data
```{r}
BIEN_traits <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_traits_forest.csv"))
TRY_traits <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_traits_forestSpecies.csv"))
GIFT_traits <- read.csv(file.path(data_path,"TropicalAndes_GIFT_plant_traits_forest.csv"))
GBIF_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ.csv"))
```

```{r}
#species list
BIEN_species <- unique(BIEN_traits$scrubbed_species_binomial)
TRY_species <- unique(TRY_traits$SpeciesName)
GIFT_species <- unique(GIFT_traits$species)
GBIF_species <- unique(GBIF_occ$species)
```

# Get higher taxonomy for species lists
```{r}
BIEN_tax<- BIEN_taxonomy_species(BIEN_species)
```

```{r}
TRY_tax <- tax_name(TRY_species, get = c('family', 'genus'), db= "itis")
```

```{r}
gift_sp <- GIFT_species()
sapply(GIFT_species, function(x) grep(x, gift_sp$work_species))
gift_sp[sapply(GIFT_species, function(x) grep(x, gift_sp$work_species)), ]

#GIFT_tax <- GIFT_taxgroup(taxon_lvl = c("family", "genus")[1], species = GIFT_species)
```

# Plot coverage
```{r}
BIEN_family_coverage <- BIEN_tax %>% count(scrubbed_family)
BIEN_family_coverage 
BIEN_genus_coverage <- BIEN_tax %>% count(scrubbed_genus) 
BIEN_genus_coverage
barplot(height = BIEN_family_coverage$n, names = BIEN_family_coverage$scrubbed_family, horiz=T, las=1)
barplot(height = BIEN_genus_coverage$n, names = BIEN_genus_coverage$scrubbed_genus, horiz=T, las=1)
```

>>>>

# Load required packages
```{r}
library(dplyr)
```

# Read in Data
```{r}
GBIF_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
GIFT_checklist <- read.csv(file.path(data_path,"TropicalAndes_plant_checklist.csv"))
```

# Plot coverage
GIFT checklist 
species 70376
genera  4381
families 273

GBIF occ
species 46,073
genera 4,885
families 581

```{r}
checklist <- c(70376, 4381, 273)
occ <- c(46073, 4885, 581)
compare_coverage <- data.frame(checklist, occ, row.names = c("species", "genera", "families"))
```

```{r}
barplot(species, df = compare_coverage)
```

### 2023-11-01 code/L2/L2_TropicalAndes_plant_traitcoverage.Rmd

---
title: "Tropical Andes plant traits"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Combinging TRY and BIEN plant trait data"
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Load required packages
```{r}
library(dplyr)
library(tidyr)
library(data.table)
```

@@

# Read in Data
```{r}
BIEN_traits <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_traits_forest.csv"))
TRY_traits <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_traits_forestSpecies.csv"))
GIFT_traits <- read.csv(file.path(data_path,"TropicalAndes_GIFT_plant_traits_forest.csv"))
```

# Subset to columns of interest
```{r}
# remove TRY columns without value in TraitID
TRY_traits_2 <- TRY_traits[!is.na(TRY_traits$TraitID), ]
```

```{r}
BIEN_traits_sub <- BIEN_traits[ , c("scrubbed_species_binomial", "trait_name", "trait_value", "unit")]
TRY_traits_sub <- TRY_traits_2[ , c("AccSpeciesName", "TraitName", "OrigValueStr", "OrigUnitStr")]
```

# Standardize column names
```{r}
BIEN_traits_stan <- BIEN_traits_sub %>% rename(species = scrubbed_species_binomial)
TRY_traits_stan <- TRY_traits_sub %>% rename(species = AccSpeciesName, trait_name = TraitName, trait_value = OrigValueStr, unit = OrigUnitStr)
```

```{r}
plant_traits <- dplyr::bind_rows(BIEN_traits_stan, TRY_traits_stan, GIFT_traits, .id = 'source')
```

```{r}
#replace source 1 or 2 with BIEN or TRY
plant_traits['source'][plant_traits['source'] == 1] <- "BIEN"
plant_traits['source'][plant_traits['source'] == 2] <- "TRY"
plant_traits['source'][plant_traits['source'] == 3] <- "GIFT"
```

# Summary
```{r}
glimpse(plant_traits)
```

```{r}
plant_traits %>%
  count(trait_name)
```

# Dispersal Traits
```{r}
# subset plant traits to just dispersal traits
plant_traits_dispersal <- plant_traits[plant_traits$trait_name == "Dispersal syndrome" | plant_traits$trait_name == "whole plant dispersal syndrome" | plant_traits$trait_name == "dispersal_syndrome",]
# look at possible trait values
plant_traits_dispersal_values <- unique(plant_traits_dispersal$trait_value)
# save list to manually classify biotic, abiotic, none, and NA
write.csv(plant_traits_dispersal_values, file.path(data_path,"plant_trait_dispersal_values.csv"))
plant_traits_dispersal %>%
  count(trait_value)
```

```{r}
# read in classified table
dispersal_classified <- read.csv(file.path(data_path,"plant_trait_dispersal_values_classified.csv"))
# reclassify dispersal trait value to table
plant_traits_dispersal$trait_value <- 
dispersal_classified$Classification[match(plant_traits_dispersal$trait_value,dispersal_classified$OrigValue )]
# check to make sure it worked
unique(plant_traits_dispersal$trait_value)
```

```{r}
# change all trait_name to dispersal_syndrome
plant_traits_dispersal['trait_name'][plant_traits_dispersal['trait_name'] == 'Dispersal syndrome'] <- 'dispersal_syndrome'
plant_traits_dispersal['trait_name'][plant_traits_dispersal['trait_name'] == 'whole plant dispersal syndrome'] <- 'dispersal_syndrome'
#need to remove none & NA
plant_traits_dispersal <- plant_traits_dispersal[plant_traits_dispersal$trait_value == "biotic" | plant_traits_dispersal$trait_value == "abiotic",]
#need one value per species
plant_traits_dispersal <- plant_traits_dispersal %>% distinct(species, .keep_all = TRUE)
plant_traits_dispersal_wide <- pivot_wider(data = plant_traits_dispersal, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = list)
```

# Plant fruit mass traits
```{r}
plant_traits_mass <- plant_traits[plant_traits$trait_name == "Fruit dry mass",]
plant_traits_mass <- plant_traits_mass[plant_traits_mass$unit == "mg",]
plant_traits_mass$trait_value <- as.numeric(plant_traits_mass$trait_value)
plant_traits_mass_wide <- pivot_wider(data = plant_traits_mass, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
```

# Plant height traits
```{r}
plant_traits_height <- plant_traits[plant_traits$trait_name == "Plant height vegetative" |plant_traits$trait_name == "whole plant height" |plant_traits$trait_name == "plant_height" ,]
plant_traits_height$trait_value <- as.numeric(plant_traits_height$trait_value)
plant_traits_height_wide <- pivot_wider(data = plant_traits_height, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
plant_traits_height_wide$plant_height <- rowMeans(plant_traits_height_wide[,2:3], na.rm = TRUE)
```

# Plant longevity traits
```{r}
plant_traits_longevity <- plant_traits[plant_traits$trait_name == "Plant lifespan (longevity)" | plant_traits$trait_name == "maximum whole plant longevity" | plant_traits$trait_name == "plant_lifespan",]
# change all trait_name to plant lifespan
plant_traits_longevity['trait_name'][plant_traits_longevity['trait_name'] == 'Plant lifespan (longevity)'] <- 'plant_lifespan'
plant_traits_longevity['trait_name'][plant_traits_longevity['trait_name'] == 'maximum whole plant longevity'] <- 'plant_lifespan'

#remove non-numeric values
plant_traits_longevity <- plant_traits_longevity[!is.na(as.numeric(as.character(plant_traits_longevity$trait_value))),]
```

```{r}
plant_traits_longevity$trait_value <- as.numeric(plant_traits_longevity$trait_value)
plant_traits_longevity_wide <- pivot_wider(data = plant_traits_longevity, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
```

# Combine all traits
```{r}
# convert all to dataframes
plant_traits_height_wide <- as.data.frame(plant_traits_height_wide)
plant_traits_dispersal_wide <- as.data.frame(plant_traits_dispersal_wide)
plant_traits_mass_wide <- as.data.frame(plant_traits_mass_wide)
plant_traits_longevity_wide <- as.data.frame(plant_traits_longevity_wide)

plant_traits_wide <- merge(x = plant_traits_height_wide, y = plant_traits_dispersal_wide, by = "species", all = TRUE)
plant_traits_wide <- merge(x = plant_traits_wide, y = plant_traits_mass_wide, by = "species", all = TRUE)
plant_traits_wide <- merge(x = plant_traits_wide, y = plant_traits_longevity_wide, by = "species", all = TRUE)
```

```{r}
plant_traits_wide <- plant_traits_wide [, c("species", "plant_height", "dispersal_syndrome", "plant_lifespan")]
# remove species with NAs for one or more traits
plant_traits_wide_complete <- na.omit(plant_traits_wide)
```

#Summary
```{r}
glimpse(plant_traits_wide_complete)
```

# Extract species list
```{r}
species_list <- unique(plant_traits_wide_complete$species)
```

```{r}
plant_traits_wide_complete$dispersal_syndrome <- as.character(plant_traits_wide_complete$dispersal_syndrome)
plant_traits_wide_complete$plant_lifespan <- as.character(plant_traits_wide_complete$plant_lifespan)
```

# Write data to csv
```{r}
write.csv(plant_traits_wide_complete, file.path(output_path,"TropicalAndes_plant_traits_forest.csv"))
write.csv(species_list, file.path(output_path,"TropicalAndes_plant_traits_forest_species_list.csv"))
```

>>>>

---
title: "Trait coverage for plant data"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script summarize trait coverage and gaps of plant data used."
data input: "ADD"
data output: "ADD"
date: "2023-10-04"
output: html_document
---

@@

# Load required packages
```{r}

```

# Read in Data
```{r}

```
